dyngen: benchmarking with \textit{in silico} single cells

### 2019-10-14 content.tex
add comment
update

@@

@@

One of the main processes involved in cellular dynamic processes is gene regulation, where regulatory cascades and feedback loops lead to progressive changes in expression and decision making. The exact way a cell chooses a certain path during its differentiation is still an active research field, although certain models have already emerged and been tested \textit{in vivo}. One driver of bifurcation seems to be mutual antagonism, where two genes \cite{xu_regulationbifurcatingcell_2015} strongly repress each other, forcing one of the two to become inactive \cite{graf_forcingcellschange_2009}. Such mutual antagonism can be modelled and simulated \cite{wang_quantifyingwaddingtonlandscape_2011, ferrell_bistabilitybifurcationswaddington_2012}. Although the two-gene model is simple and elegant, the reality is frequently more complex, with multiple genes (grouped into modules) repressing each other \cite{yosef_dynamicregulatorynetwork_2013}.

@@

\subsection{Convert gene regulatory network to a set of reactions} \label{sec:reactions}
\newcommand{\w}[1]{\text{w}_{#1}}
\newcommand{\x}[1]{\text{x}_{#1}}
\newcommand{\y}[1]{\text{y}_{#1}}

>>>>

% YSA: RNA velocity: Onduidelijk hier, mensen kennen dit niet.

@@

% We developed \texttt{dyngen}, a simulator of single cells to benchmark and stress-test computational tools for single-cell omics.
%It supports a selection of existing experimental protocols, but can also simulate experiments for permitted by current technologies, such as generating omics profiles from the same cell at multiple timepoints.
%This data can help kick-start emerging domains with low data availability more safely by allowing method developers to test their method before numerous datasets become publicly available.
%\texttt{dyngen} has already been used successfully to evaluate trajectory inference\cite{saelens_comparisonsinglecelltrajectory_2019}, trajectory alignment\cite{vandenberge_trajectorybaseddifferentialexpression_2019}, and network inference\cite{pratapa_benchmarkingalgorithmsgene_2019} methods. Furthermore, we showed that it can also be used to evaluate differential network inference methods, and trajectory alignment methods. % TODO: And RNA velocity?

@@

One of the main processes involved in cellular dynamic processes is gene regulation, where regulatory cascades and feedback loops lead to progressive changes in expression and decision making. The exact way a cell chooses a certain path during its differentiation is still an active research field, although certain models have already emerged and been tested \textit{in vivo}. One driver of bifurcation seems to be mutual antagonism, where two genes strongly repress each other\cite{rekhtman_directinteractionhematopoietic_1999,xu_regulationbifurcatingcell_2015}, forcing one of the two to become inactive \cite{graf_forcingcellschange_2009}. Such mutual antagonism can be modelled and simulated \cite{wang_quantifyingwaddingtonlandscape_2011, ferrell_bistabilitybifurcationswaddington_2012}. Although the two-gene model is simple and elegant, the reality is frequently more complex, with multiple genes (grouped into modules) repressing each other \cite{yosef_dynamicregulatorynetwork_2013}.

@@

\subsection{Convert gene regulatory network to a set of reactions} \label{sec:reactions}
% paper on transcriptional bursting
% doi: 10.1039/C7MB00154A
% interesting to try to replicate in dyngen! :)
\newcommand{\w}[1]{\text{w}_{#1}}
\newcommand{\x}[1]{\text{x}_{#1}}
\newcommand{\y}[1]{\text{y}_{#1}}
